Discrepancy, Coresets, and Sketches in Machine Learning

### 2019-01-31 komlos.tex
post manual merge

Logistic Regression is an extremely popular technique for classification. We are given a set of $n$ pairs $x_i,y_i$ with $x_i \in \R^d$ and $y_i \in \{-1,1\}$. The objective is to find a regressor $q \in \R^d$ minimizing the logistic loss defined as $\E_i \log(1+\exp(x_i^\top q y_i))$. A coreset for logistic regression is a set of $m \ll n$ pairs $z_i, v_i$ in $\R^d, \{-1,1\}$ correspondingly along with weights $w_i$, with the property that for any regressor $q$, either $\E_i \log(1+\exp(x_i^\top q y_i)) = \E_i \log(1+\exp(z_i^\top q v_i)) \pm \eps$ or $\E_i \log(1+\exp(x_i^\top q y_i)) = \E_i \log(1+\exp(z_i^\top q v_i)) (1 \pm \eps)$. A recent paper \cite{DBLP:journals/corr/abs-1805-08571} provides a coreset with a multiplicative guarantee that is based on an average sensitivity property of the dataset. They provide a lower bound for the size of a multiplicative coreset showing that in general it is not possible to achieve $m \ll n$. The coreset they build is of cardinality $m \approx \mu\sqrt{nd^3}/\eps^2$ where $\mu \geq 1$ is complexity measure of the dataset.  \cite{tolochinsky2018coresets} gives a generic multiplicative coreset construction for any monotonic function with $\ell_2^2$ regularization. The dependence they get is $\tilde O(d/\eps^2)$ ignoring logarithmic factors.  Our results mentioned above prove that for an additive guarantee, it is possible to obtain a coreset of size $O(\sqrt{d}/\eps)$.

@@

\begin{definition}[\bf Class Discrepancy] \label{cd1}
Let $A \subset \R^m$ and $\sigma \in \{-1,1\}^m$ the class discrepancy of $A$ is 
$
K_m(A)  = \min_\sigma \max_{a\in A} \left| \frac{1}{m} \sum_{i=1}^{m}  \sigma_i  a_i\right| 
$.
\end{definition}
%
\begin{definition}[\bf Class Discrepancy]\label{cd2}
Let $f:\X,\Q\rightarrow\R$ and $\sigma \in \{-1,1\}^m$. 
The class discrepancy of $f$ w.r.t.\ $\{x_1,\ldots,x_m\} \subset \X$ is 
$
K_m(f) =  \min_{\sigma} \max_{q \in \Q}  \left|\frac{1}{m} \sum_{i=1}^{m}\sigma_i f(x_i,q)\right|
$.
\end{definition}
%
\begin{definition}[\bf Class Discrepancy]\label{cd3}
Let $\F$ be a family of functions $f:\X\rightarrow\R$ and $\sigma \in \{-1,1\}^m$. 
The class discrepancy of $\F$ w.r.t.\ $\{x_1,\ldots,x_m\} \subset \X$ is 
$
K_m(\F) =  \min_{\sigma} \max_{f \in \F}  \left| \frac{1}{m} \sum_{i=1}^{m}\sigma_i f(x_i)\right|
$.
\end{definition}
The class discrepancy of $f$ or $\F$ without a reference a set $\{x_1,\ldots,x_m\}$ is the upper bound on any subset of $\X$ of size $m$.
Throughout the manuscript we assume a bijective mapping between $\F$ and $\Q$. 
Specifically, $f \in \F$ if there exists $q \in \Q$ such that $f(x) = f(x,q)$ and vice versa. 
In the context of machine learning, one should think about $f(x,q)$ as the loss associated with example $x$ and model parameters $q$.
Definitions~\ref{cd2} and \ref{cd3} are therefore equivalent. 
For the equivalence of Definition~\ref{cd1}, the set $A$ should be thought of as the set of possible induces losses. 
Namely, $a\in A$ if there is a model $q$ such $a_i = f(x_i,q)$.
The three equivalent definitions between the definitions is therefore straight forward.

These definition are very similar to the Rademacher complexity $R_m(\F)$.
The only difference is that we consider the minimal value over $\sigma$ rather than its expectation for a uniform random $\sigma$.
Rademacher complexity uses random signs because it measures the effect of data being chosen uniformly at random (or drawn from an unknown distribution). See \cite{Bartlett:2003:RGC:944919.944944} for other notions of complexity and relationships between them.\footnote{\cite{Bartlett:2003:RGC:944919.944944} use the name ``discrepancy'' to define a different quantity than the one above.}
 We claim that $K_m$ is the measure more suitable for creating coresets and for creating mergeable sketches.
Trivially,  $K_m \le R_m$. We show, however, that $K_m(\F) = o(R_m(\F))$ for a wide range of interesting problems in machine learning.

>>>>

Logistic Regression is an extremely popular technique for classification. We are given a set of $n$ pairs $x_i,y_i$ with $x_i \in \R^d$ and $y_i \in \{-1,1\}$. The objective is to find a regressor $q \in \R^d$ minimizing the logistic loss defined as $\E_i \log(1+\exp(x_i^\top q y_i))$. A coreset for logistic regression is a set of $m \ll n$ pairs $z_i, v_i$ in $\R^d, \{-1,1\}$ correspondingly along with weights $w_i$, with the property that for any regressor $q$, either $\E_i \log(1+\exp(x_i^\top q y_i)) = \E_i \log(1+\exp(z_i^\top q v_i)) \pm \eps$ or $\E_i \log(1+\exp(x_i^\top q y_i)) = \E_i \log(1+\exp(z_i^\top q v_i)) (1 \pm \eps)$. A recent paper \cite{DBLP:journals/corr/abs-1805-08571} provides a coreset with a multiplicative guarantee that is based on an average sensitivity property of the dataset. They provide a lower bound for the size of a multiplicative coreset showing that in general it is not possible to achieve $m \ll n$. The coreset they build is of cardinality $m \approx \mu\sqrt{nd^3}/\eps^2$ where $\mu \geq 1$ is a complexity measure of the dataset.  \cite{tolochinsky2018coresets} gives a generic multiplicative coreset construction for any monotonic function with $\ell_2^2$ regularization. The dependence they get is $\tilde O(d/\eps^2)$ ignoring logarithmic factors.  Our results prove that for an additive guarantee, it is possible to obtain a coreset of size $O(\sqrt{d}/\eps)$.

@@

\begin{definition}[\bf Class Discrepancy] \label{cd1}
Let $A \subset \R^m$ and $\sigma \in \{-1,1\}^m$ the class discrepancy of $A$ is 
$
K_m(A)  = \min_\sigma \max_{a\in A} \left| \frac{1}{m} \sum_{i=1}^{m}  \sigma_i  a_i\right| 
$.
\end{definition}
%
\begin{definition}[\bf Class Discrepancy]\label{cd2}
Let $f:\X,\Q\rightarrow\R$ and $\sigma \in \{-1,1\}^m$. 
The class discrepancy of $f$ w.r.t.\ $\{x_1,\ldots,x_m\} \subset \X$ is 
$
K_m(f) =  \min_{\sigma} \max_{q \in \Q}  \left|\frac{1}{m} \sum_{i=1}^{m}\sigma_i f(x_i,q)\right|
$.
\end{definition}
%
\begin{definition}[\bf Class Discrepancy]\label{cd3}
Let $\F$ be a family of functions $f:\X\rightarrow\R$ and $\sigma \in \{-1,1\}^m$. 
The class discrepancy of $\F$ w.r.t.\ $\{x_1,\ldots,x_m\} \subset \X$ is 
$
K_m(\F) =  \min_{\sigma} \max_{f \in \F}  \left| \frac{1}{m} \sum_{i=1}^{m}\sigma_i f(x_i)\right|
$.
\end{definition}
The class discrepancy of $f$ or $\F$ without a reference a set $\{x_1,\ldots,x_m\}$ is the upper bound on any subset of $\X$ of size $m$.
Throughout the manuscript we assume a bijective mapping between $\F$ and $\Q$. 
Specifically, $f \in \F$ if there exists $q \in \Q$ such that $f(x) = f(x,q)$ and vice versa. 
In the context of machine learning, one should think about $f(x,q)$ as the loss associated with example $x$ and model parameters $q$.
%Definitions~\ref{cd2} and \ref{cd3} are therefore equivalent. 
The set $A$ should be thought of as the set of all possible induces losses vectors. 
Namely, $a\in A$ if there is a model $q$ such $a_i = f(x_i,q)$.

\todo{This paragraph seems to be repeated/duplicated bellow.}

These definition are intentionally very similar to those of the Rademacher complexity $R_m(\F)$.
The only difference is that we consider the minimal value over $\sigma$ rather than its expectation for a uniform random $\sigma$.
Rademacher complexity uses random signs because it measures the effect of data being chosen uniformly at random (or drawn from an unknown distribution). See \cite{Bartlett:2003:RGC:944919.944944} for other notions of complexity and relationships between them.\footnote{\cite{Bartlett:2003:RGC:944919.944944} use the term ``discrepancy'' to define a different quantity than the one above.}
We claim that $K_m$ is the right measure of complexity for bounding the size of coresets (coreset complexity) and for creating mergeable sketches (streaming coreset complexity).
Trivially,  $K_m \le R_m$. 
We show, however, that $K_m(\F) = o(R_m(\F))$ for a wide range of interesting problems in machine learning.
This immediately proves the existence of small coresets for these problems.
